---
author: "Garima Setia"
date: "9/28/2021"
output:
  html_document
---






```{r}
A = matrix(c(0, 0, 1.73, 0.50, 0.95, 0.60, 0.03, 0.03,  0.38), nrow = 3, byrow = TRUE)
A

eigen.A = eigen(A)
eigen.A

#Dominate eigenvalue
dom.pos = which.max(Re(eigen.A$values))
L1 = Re(eigen.A$values[dom.pos])
L1
# dominate eigenvalue = 1.023189

#Stable-stage/age structure

w = Re(eigen.A$vectors[,dom.pos])

SSD = w/sum(w)
round(SSD, 3)
#The population of Mariposa Lily will contain 7.5% seeds, 88% non-flowering rosettes and 4.5% flowering rosettes.

#Reproductive value
M = eigen(t(A)) #transpose of the projection matrix A
v = Re(M$vectors[, which.max(Re(M$values))])

RV = v/v[1]
round(RV, 3)

#Along with increase in age, the reproductive value is also increasing (1, 1.785, 4.355). 

vw.s = v %*% t(w)
S = vw.s/as.numeric(v%*%w)
S
round(S,3)
## s_{32} = 2.082 is the most sensitive element in the transition matrix

```
# dominate eigenvalue = 1.023189
#The population of Mariposa Lily will contain 7.5% seeds, 88% non-flowering rosettes and 4.5% flowering rosettes.
#Along with increase in age, the reproductive value is also increasing (1, 1.785, 4.355). 
# s_{32} = 2.082 is the most sensitive element in the transition matrix



```{r}

B = matrix(c(0, 0, 1.73, 0.50, 0.875, 0.60, 0.03, 0.03, 0.38), nrow = 3, byrow = TRUE)
B

N0 = matrix(c(200, 200, 200), ncol = 1)
N0

N1 = B%*%N0
N1

years = 50
N = matrix(0, ncol = length(N0), nrow = years + 1)
N[1,] = N0
Simulations = 1000
Total.population = numeric()
Df.extinction<-as.data.frame(matrix(nrow = years, ncol = Simulations))

for(j in 1:Simulations){
  N = matrix(0, ncol = length(N0), nrow = years + 1)
  N[1,] = N0
  for(i in 1:years){
    YearMatrix<-sample(x = c('Good','Bad'), size = 1, prob = c(.5,.5))
    if(YearMatrix=='Bad'){
      popt<-B
    }
    if(YearMatrix=='Good'){
      popt<-A
    }
    N[i+1,] = popt%*%N[i,]
    if(sum(N[i+1,])>100){
      Df.extinction[i,j]<-0
    }
    if(sum(N[i+1,])<=100){
      Df.extinction[i,j]<-1
    }
  }
  Total.population[j]<-sum(N[i,])
}

```
#Probability of extinction

```{r}
hist(Total.population, xlim = c(0, max(Total.population)))
abline(v = 100, col='red')
legend('topright', legend = 'Quasi-extinction level',
       col = 'red', bty='n',lwd=1)


probOfExtinction<-numeric()
time<-1:50

for(i in 1:nrow(Df.extinction)){
  probOfExtinction[i]<-sum(Df.extinction[i,])/Simulations
}

plot(x = time, y = probOfExtinction, main = 'Probability of extinction', xlab = 'years', ylab = 'Probability', type = 'l')
```


#Since the quasi-extinction level (100) is less than the lowest population level and the value obtained for the probability of extinction is zero, it can be said that population is not on a verge of extinction.
